Fixing Vulnerabilities Hinders Maintainability

### 2020-01-27 paper/icpc20.tex
charts updated

\renewcommand{\shortauthors}{Reis et al.}

@@

\begin{itemize}
    \item \textbf{CW-89: Improper Neutralization of Special Elements used in an SQL Command/
	SQL Injection (23 cases).} When developers do not keep untrusted data
    separate from SQL queries. If an attacker sends a command that
    exploits the syntax of the SQL interpreter, then SQL injection attack is possible.

@@

    \item \textbf{CWE-20: Improper Input Validation (75 cases).} When developers
	forget to validate the inputs of an application, attackers may have control
	of the control/data flow of the program.

    \item \textbf{CWE-200: Information Exposure (38 cases).}

	Application's logs are one way of intentional/unintentional disclosure information
	to attackers. Many times attackers get access to logs when sould not be authorized to.

	\item \textbf{CWE-352: Cross-Site Request Forgery (CSRF) (20 cases).} Poor session tokens generation
    and management usually allow attackers to send forged HTTP requests
    including authentication information from the victim to the vulnerable web
    application.

	\item \textbf{CWE-264: Permissions, Privileges, and Access Controls (36 cases).}

	Incorrectly
	    implemented functionalities related to authentication and session
	    management, allowing an attacker to gain access to session tokens,
	    passwords, keys, and other sensitive data.

	\item \textbf{CWE-399: Resource Management Errors (20 cases).} 
	Resource management issues are found more frequently
		in programming languages that do not manage memory automatically (e.g., C/C++
	    and Objective-C), i.e., where developers are responsible for
	    handling it instead. It is one of the main causes of DoS attacks.

@@

	\item \textbf{CWE-22: Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (36 cases).} 
	When developers forget to neutralize special elements within the pathnames to files/directories, attackers 
	may leverage this flaw to resolve to a location outside of the restricted directory and gain access to documents
	or entire repositories.

@@

\begin{figure}[h]
	\vspace{-0.2cm}

 	\centering 	\includegraphics[width=0.46\textwidth]{figures/maintainability.pdf}
 	\caption{Maintainability difference for security and baseline refactorings}
	\label{fig:secvsreg}
	\vspace{-0.2cm}

@@

\begin{figure}[h]
\vspace{-0.2cm}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/language.pdf}
  \caption{Maintainability difference by language after security refactorings.}
  \label{fig:lang_main}
  \vspace{-0.2cm}

@@

\begin{figure}[h]
	\vspace{-0.2cm}
 	\centering
 	\includegraphics[width=0.57\textwidth]{figures/category.pdf}
 	\caption{Maintainability difference by pattern after security refactorings.}
	\label{fig:pat}
	\vspace{-0.2cm}

>>>>

\renewcommand{\shortauthors}{Anon. et al.}

@@

\begin{itemize}
    \item \textbf{CW-89: Improper Neutralization of Special Elements used in an SQL Command/SQL Injection (23 cases).} When developers do not keep untrusted data
    separate from SQL queries. If an attacker sends a command that
    exploits the syntax of the SQL interpreter, then SQL injection attack is possible.

@@

    \item \textbf{CWE-20: Improper Input Validation (76 cases).} When developers
	forget to validate the inputs of an application, attackers may have control
	of the control/data flow of the program.

    \item \textbf{CWE-200: Information Exposure (39 cases).} Application's logs are one way of intentional/unintentional disclosure information
	to attackers. Many times attackers get access to logs when sould not be authorized to.

	\item \textbf{CWE-352: Cross-Site Request Forgery (CSRF) (22 cases).} Poor session tokens generation
    and management usually allow attackers to send forged HTTP requests
    including authentication information from the victim to the vulnerable web
    application.

	\item \textbf{CWE-264: Permissions, Privileges, and Access Controls (59 cases).} Incorrectly
	    implemented functionalities related to authentication and session
	    management, allowing an attacker to gain access to session tokens,
	    passwords, keys, and other sensitive data.

	\item \textbf{CWE-399: Resource Management Errors (22 cases).} Resource management issues are found more frequently
		in programming languages that do not manage memory automatically (e.g., C/C++
	    and Objective-C), i.e., where developers are responsible for
	    handling it instead. It is one of the main causes of DoS attacks.

@@

	\item \textbf{CWE-22: Improper Limitation of a Pathname to a Restricted Directory/Path Traversal (37 cases).}  When developers forget to neutralize special elements within the pathnames to files/directories, attackers 
	may leverage this flaw to resolve to a location outside of the restricted directory and gain access to documents
	or entire repositories.

@@

\begin{figure}[h]	
 	\centering 	\includegraphics[width=0.46\textwidth]{figures/maintainability_sec_vs_reg.pdf}
 	\caption{Maintainability difference for security and baseline patches}
	\label{fig:secvsreg}

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.57\textwidth]{figures/impact_per_guideline.pdf}
 	\caption{Impact of a security patch in 8 out of the 10 guidelines.}
	\label{fig:guidelines}

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.5\textwidth]{figures/maintainability_severity.pdf}
 	\caption{Maintainability difference by the vulnerabilities severity.}
	\label{fig:severity}

@@

\begin{figure}[h]
\vspace{-0.2cm}
  \centering
  \includegraphics[width=0.5\textwidth]{figures/maintainability_language.pdf}
  \caption{Maintainability difference by language after security refactorings.}
  \label{fig:lang_main}

@@

\begin{figure}[h]
 	\centering
 	\includegraphics[width=0.5\textwidth]{figures/maintainability_cwe.pdf}
 	\caption{Maintainability difference by pattern after security patches.}
	\label{fig:pat}
